---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tilt.company.match

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![tilt.company.match status badge](https://2degreesinvesting.r-universe.dev/badges/tilt.company.match)](https://2degreesinvesting.r-universe.dev)
<!-- badges: end -->

The goal of tilt.company.match is to provide helpers for company name
matching in the tilt-project.

## Installation

You can install the development version of
[tilt.company.match](https://github.com/2DegreesInvesting/tilt.company.match)
from [r-universe](https://r-universe.dev/) with:

```r
options(repos = c("https://2degreesinvesting.r-universe.dev", getOption("repos")))
install.packages("tilt.company.match")
```

Or you can install it from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/tilt.company.match")
```

## Example

Here is a minimal example of what you can do with the package
tilt.copmany.match. For a complete and gentle walk-through see [Get
started](https://2degreesinvesting.github.io/tilt.company.match/articles/tilt-company-match.html).

```{r}
library(vroom, warn.conflicts = FALSE)
library(tilt.company.match)

# TODO: Replace with the path/to/your/real/loanbook.csv
loanbook_csv <- example_file("demo_loanbook.csv")
loanbook_csv

loanbook <- vroom(loanbook_csv, show_col_types = FALSE)
loanbook

# TODO: Replace with the path/to/your/real/tilt.csv
tilt_csv <- example_file("demo_tilt.csv")
tilt_csv

tilt <- vroom(tilt_csv, show_col_types = FALSE)
tilt

check_loanbook(loanbook)

suggest_match(loanbook, tilt)
```

[Get
started](https://2degreesinvesting.github.io/tilt.company.match/articles/tilt-company-match.html).
