---
title: "Calculate coverage statistics"
author: "Kalash Singhal"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

Use the required packages, read your `loanbook` and `tilt` datasets, check your
`loanbook` and suggest matches. For a gentle walkthrough see [Get
started](https://2degreesinvesting.github.io/tilt.company.match/articles/tilt-company-match.html).

```{r setup}
library(dplyr, warn.conflicts = FALSE)
library(vroom)
library(tilt.company.match)

loanbook <- vroom(example_file("demo_loanbook.csv"), show_col_types = FALSE)
tilt <- vroom(example_file("demo_tilt.csv"), show_col_types = FALSE)

loanbook %>% check_loanbook()

to_edit <- loanbook %>% suggest_match(tilt)

# Edited file after manual validation (File replacement from the user is required)
edited <- vroom(example_file("demo_matched.csv"), show_col_types = FALSE)
edited
```

## Share of matched companies after manual validation from total loanbook companies

```{r}
accepted <- edited %>%
  filter(accept_match)

matched_rows <- accepted %>% nrow()

total_rows <- loanbook %>%
  distinct(id) %>%
  nrow()

matched_share <- matched_rows / total_rows
matched_share
```

## Number and share of matched companies classified by `misc_info` from total loanbook companies

Please replace the `misc_info` column with variables like sectors, headcount or
similar classifier to calculate the share based on that classifier.

```{r}
x_misc <- loanbook %>% count(misc_info)
y_misc <- accepted %>% count(misc_info)
misc_share <-
  left_join(x_misc, y_misc, by = c("misc_info"), suffix = c("_total", "_merged")) %>%
  mutate(n_share = n_merged / n_total)
```

## Sum of loan values and their share from total value after classified by `misc_info`

The `loan_amount` column is added as a sample column to calculate the loan
exposure output.

```{r}
# Sample column and dataframe
accepted_loan <- accepted %>% mutate(loan_amount = c(1000))

# Please replace the name of sample column `loan_amount` and the `accepted_loan`
# dataframe with the original loan column name and loanbook dataframe!
exposure_sum <- accepted_loan %>%
  group_by(misc_info) %>%
  summarise(sum_expo = sum(loan_amount)) %>%
  mutate(sum_share = sum_expo / sum(sum_expo))
```
