---
title: "Handling a large loanbook"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

If your `loanbook` is large or you lack `postcode` or `country` your computer
may run out of memory and crash. This article shows an alternative way to run
the first step of the matching process: Suggesting matches. You'll check the
quality of your `loanbook` before expensive computations, and match multiple
small chunks of your `loanbook` one by one (like
[here](https://rmi-pacta.github.io/r2dii.match/articles/chunk-your-data.html)).

```{r}
library(dplyr, warn.conflicts = FALSE)
library(vroom)
library(fs)
library(tilt.company.match)
```

Read your `loanbook` and `tilt` datasets.

```{r}
# TODO: Replace with the /path/to/your/real/tilt.csv
tilt <- vroom(example_file("demo_tilt.csv"), show_col_types = FALSE)

# TODO: Replace with the /path/to/your/real/loanbook.csv
loanbook <- vroom(example_file("demo_loanbook.csv"), show_col_types = FALSE)
```

Check and fix your data first -- before engaging in expensive computations.

```{r}
loanbook %>% check_loanbook()
```

The columns `postcode` and `country` will save you time but aren't crucial. If
you don't have them, fake them and move on. For example, if you lack `postcode`
fake it with `mutate(loanbook, postcode = NA)`.

More chunks take longer to run but use less memory; you'll need to experiment to
find the number of chunks that best works for you.

```{r}
chunks <- 3
chunked <- loanbook %>% mutate(chunk = as.integer(cut(row_number(), chunks)))

# This "output" directory is temporary; you may use any folder in your computer
out <- path(tempdir(), "output")
dir_create(out)

# In your console this shows a progress bar
pb <- progress::progress_bar$new(total = length(unique(chunked$chunk)))
for (i in unique(chunked$chunk)) {
  pb$tick()

  # 1. Match this chunk against the entire `tilt` dataset.
  this_chunk <- filter(chunked, chunk == i)
  this_result <- suggest_match(this_chunk, tilt)

  # 2. If this chunk matched nothing, move to the next chunk
  matched_nothing <- nrow(this_result) == 0L
  if (matched_nothing) next()

  # 3. Else, save the result to a .csv file.
  vroom_write(this_result, path(out, paste0(i, ".csv")))
}
```

Read the output of each chunk into a single dataframe.

```{r}
to_edit <- vroom(dir_ls(out))
to_edit
```

Write it back to a single .csv and continue with steps 2 and 3.

```{r}
to_edit %>% vroom_write("to_edit.csv")
```



